Andres Karjus &nbsp;<font size='3'> evolutionary&nbsp;linguist | cultural&nbsp;data&nbsp;analyst | AI&nbsp;&&nbsp;DH&nbsp;instructor </font>

### 2023-11-12 index.Rmd
news

* Andres Karjus. 2023. Machine-assisted mixed methods: augmenting humanities and social sciences with artificial intelligence | [preprint](https://arxiv.org/abs/2309.14379) | [data&code](https://github.com/andreskarjus/MachineAssistedMixedMethods)
* Andres Karjus and Christine Cuskley. 2023. Evolving Linguistic Divergence on Polarizing Social Media | arXiv [preprint](https://arxiv.org/abs/2309.01659)
* Tillmann Ohm, Mar Canet Solà, Andres Karjus, Maximilian Schich. 2023. Collection Space Navigator: An Interactive Visualization Interface for Multidimensional Datasets. VINCI 2023: Proceedings of the 16th International Symposium on Visual Information Communication and Interaction. | [open access](doi.acm.org?doi=3615522.3615546) | [extended preprint](https://arxiv.org/abs/2305.06809) | [interactive demo](https://collection-space-navigator.github.io/) | [code](https://github.com/Collection-Space-Navigator/CSN)
* Andres Karjus, Mar Canet Solà, Tillmann Ohm, Sebastian Ahnert, Maximilian Schich. 2023. Compression ensembles quantify aesthetic complexity and the evolution of visual art. EPJ Data Science. 12, 21. | [open access](https://doi.org/10.1140/epjds/s13688-023-00397-3) | [code](https://github.com/andreskarjus/compression_ensembles)
* Oiva, Mila, Ksenia Mukhina, Vejune Zemaityte, Tillmann Ohm, Mikhail Tamm, Andres Karjus, Mark Mets, Daniel Chávez Heras Mar Canet Solà, Helena Hanna Juht, Maximilian Schich. 2023. A Framework for the Analysis of Historical Newsreels | [preprint](https://doi.org/10.31235/osf.io/a4xsp)
* Mehmet Burak Yilmaz, Elen Lotman, Andres Karjus, Pia Tikka. 2023. An embodiment of the cinematographer: emotional and perceptual responses to different camera movement techniques. Frontiers Neuroscience | [open access](https://doi.org/10.3389/fnins.2023.1160843) | [data](https://zenodo.org/record/7979713)
* Juan Guerrero Montero, Andres Karjus, Kenny Smith, Richard A. Blythe. 2023. Reliable identification of selection mechanisms in language change | [preprint](https://arxiv.org/abs/2305.15914)
* Vejune Zemaityte, Andres Karjus, Ulrike Rohn, Maximilian Schich, Indrek Ibrus. 2023. Quantifying the global film festival circuit: Networks, diversity, and public value creation | [preprint](https://doi.org/10.31235/osf.io/g9w4b) | [code & data](https://andreskarjus.github.io/cinandofestivals)
* Indrek Ibrus, Andres Karjus, Vejune Zemaityte, Ulrike Rohn, Maximilian Schich. 2022. Quantifying public value creation by public service media using big programming data | [preprint](https://osf.io/preprints/socarxiv/hysma/)
* Mark Mets, Andres Karjus, Indrek Ibrus, Maximilian Schich. 2023. Automated stance detection in complex topics and small languages: the challenging case of immigration in polarizing news media | [preprint](https://arxiv.org/abs/2305.13047)
* Andres Karjus, Christine Cuskley. 2022. Evolving linguistic divergence in socio-political polarities. Proceedings of the Joint Conference on Language Evolution (JCoLE). | [proceedings pdf](http://www.evolang.org/jcole2022_proceedings/jcole2022_proceedings.pdf) 

@@

* Andres Karjus, 2020. Competition, selection and communicative need in language change. PhD thesis, University of Edinburgh | [pdf](https://andreskarjus.github.io/thesis/Andres_Karjus_phd_thesis.pdf) | [1-page non-technical summary](https://andreskarjus.github.io/thesis/english_summary.pdf) | [eestikeelne lühikokkuvõte](https://andreskarjus.github.io/thesis/doktoritoo_ylevaade_AK.pdf)
*  Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Communicative need modulates competition in language change | [preprint](https://arxiv.org/abs/2006.09277)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Challenges in detecting evolutionary forces in language change using diachronic corpora. Glossa: a journal of general linguistics, 5(1), p.45. | [open access](http://doi.org/10.5334/gjgl.909) | [code](https://github.com/andreskarjus/wfsim_fit)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2020. Quantifying the dynamics of topical fluctuations in language. Language Dynamics and Change 10(1), 86-125 | [open access](https://doi.org/10.1163/22105832-01001200) | [code](https://github.com/andreskarjus/topical_cultural_advection_model)
* Andres Karjus, Martin Ehala, 2018. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change, 8, pp. 219-252 | [journal link](https://doi.org/10.1163/22105832-00802004) | [open postprint](Karjus_Ehala_2018_Testing_an_agent-based_model_of_language_choice_LDCpostprint.pdf) | <a href="karjus_testing_2018.bib" target="_blank">bib</a> | sociolinguistic [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): 1000 respondents, 200 questions
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract, full paper above] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>

@@

- Talk on LLMs for studying texts and change at "Assessing and measuring systems change" workshop | [slides](https://andreskarjus.github.io/talks/Tartu_LLMtalk.pdf)
- Poster on film festival research at Netsci 2023, Vienna
- Presentation at the Bibliotheca Herziana workshop on computational approaches to art
- Talk "Programming, data visualization & AI  for academic audiences across institutions and disciplines: lessons learned", at the Cross-university collaboration in Digital Humanities & Social Science (DHSS) and Digital Humanities & Cultural Heritage (DHCH) Education workshop of the DHNB2023 conference | [slides](https://datafigure.eu/reports/AndresKarjusDatafigure_CrossUniversityDHworkshop.pdf)
- Talk "Exploring Estonian Public Television Production 2004-2020 Using Big Programming Data" at the 8th Estonian Digital Humanities [Conference](https://dh.org.ee/events/dhe2022) (05.10.2022)
- Poster "Evolving Linguistic Divergence in Socio-Political Polarities", at the [JCoLe](https://sites.google.com/view/joint-conf-language-evolution) Joint Conference on Language Evolution in Kanazawa, Japan (August 2022)
- Poster "Linguistic divergence in American English along socio-political polarities", the [IC2S2](https://www.ic2s2.org/) Computational Social Science Conference (20.07.2022) | [pdf](https://andreskarjus.github.io/talks/ic2s2/IC2S2_poster_Linguistic_divergence_in_American_English_along_socio-political_polarities.pdf)
- Seminar talk at the Poncelet laboratory in Moscow (November 2021)
- [Conference on Complex Systems 2021](https://ccs2021.univ-lyon1.fr/) (October 2021) | [slides](https://andreskarjus.github.io/talks/ccs2021/Karjus_CCS2021_Capturing_aesthetic_complexity.pdf)
- [Protolang 7](https://blogs.phil.hhu.de/protolang7/programme/) (September 2021) | [slides](https://andreskarjus.github.io/talks/andres_karjus_protolang7.pdf)
- [Culture Conference 2021](https://culture-conference.com/) | [Poster on aesthetic complexity](https://andreskarjus.github.io/talks/cultureconference2021/)
- [TÜling](https://www.keel.ut.ee/et/tuling) (April 2021)
- [Colloquium for Computational Linguistics and Linguistics in Stuttgart](https://www.ims.uni-stuttgart.de/en/research/colloquium/) | [Slides](https://andreskarjus.github.io/talks/stuttgart2021/andres_karjus_lexical_dynamics_slides_stuttgart2021.pdf) | [Recording](https://www.ims.uni-stuttgart.de/documents/aktuelles/veranstaltungen/210111-Karjus-Talk-Colloq-Recording.mp4)
- [RUSE 2019](http://rusesymposium.org.uk/). Slides [here](https://andreskarjus.github.io/ruse2019).
- [CL2019](http://www.cl2019.org/). Slides [here](https://andreskarjus.github.io/cl2019talk.html).
- [Culture Conference 2019](https://culture-conference.com/programme-2019/), Poster [here](https://andreskarjus.github.io/lexcom_poster).
- [Inaugural ISLE workshop](https://www.comparativelinguistics.uzh.ch/en/events/ISLE-inaugural-workshop.html) | _Modelling lexical interactions in diachronic corpora_ |  [poster](https://andreskarjus.github.io/isle_poster)
- University of Edinburgh [Centre for Language Evolution](http://www.lel.ed.ac.uk/cle/) seminar series | _Challenges in detecting evolutionary forces in language change using diachronic corpora_ | [slides](https://andreskarjus.github.io/evoforces_cletalk/slides.html) | [code](https://github.com/andreskarjus/wfsim_fit)
- Corpus Linguistics in Scotland Network Meeting, _Topical Fluctuations and Lexical Interactions in Diachronic Corpora_

@@

I also teach workshops as an instructor in the private sector. 
Some of these have been standalone events, some have been part of conferences, summer schools or academic retreats. 
Feel free to get in touch if you are interested in talking about organizing a workshop on anything related to data science and statistics, artificial intelligence, data visualization, R, corpus linguistics, digital humanities, etc. <br>
For more details, head over to [datafigure.eu](https://datafigure.eu)

@@

Academic email: andres.karjus --at-- tlu.ee <br>
Business/workshop inquiries: please contact via the Datafigure page 
<br>
<br>
Twitter / X: <a href=https://twitter.com/AndresKarjus>twitter.com/AndresKarjus</a> <br>
Bluesky: <a href=https://bsky.app/profile/andreskarjus.bsky.social>andreskarjus.bsky.social</a> <br>
Mastodon: <a href=https://mastodon.social/@andreskarjus>mastodon.social/@AndresKarjus</a>
<br>
LinkedIn: <a href="https://www.linkedin.com/in/andreskarjus/">www.linkedin.com/in/andreskarjus/</a> <br> 
</div>

>>>>

2023

* Andres Karjus. 2023. Machine-assisted mixed methods: augmenting humanities and social sciences with artificial intelligence | [preprint](https://arxiv.org/abs/2309.14379) | [data&code](https://github.com/andreskarjus/MachineAssistedMixedMethods)
* Juan Guerrero Montero, Andres Karjus, Kenny Smith, Richard A. Blythe. 2023. Reliable identification of selection mechanisms in language change. Corpus Linguistics and Linguistic Theory | [Open access](https://doi.org/10.1515/cllt-2023-0064)
* Andres Karjus and Christine Cuskley. 2023. Evolving Linguistic Divergence on Polarizing Social Media | arXiv [preprint](https://arxiv.org/abs/2309.01659)
* Tillmann Ohm, Mar Canet Solà, Andres Karjus, Maximilian Schich. 2023. Collection Space Navigator: An Interactive Visualization Interface for Multidimensional Datasets. VINCI 2023: Proceedings of the 16th International Symposium on Visual Information Communication and Interaction. | [open access](doi.acm.org?doi=3615522.3615546) | [extended preprint](https://arxiv.org/abs/2305.06809) | [interactive demo](https://collection-space-navigator.github.io/) | [code](https://github.com/Collection-Space-Navigator/CSN)
* Andres Karjus, Mar Canet Solà, Tillmann Ohm, Sebastian Ahnert, Maximilian Schich. 2023. Compression ensembles quantify aesthetic complexity and the evolution of visual art. EPJ Data Science. 12, 21. | [open access](https://doi.org/10.1140/epjds/s13688-023-00397-3) | [code](https://github.com/andreskarjus/compression_ensembles)
* Oiva, Mila, Ksenia Mukhina, Vejune Zemaityte, Tillmann Ohm, Mikhail Tamm, Andres Karjus, Mark Mets, Daniel Chávez Heras Mar Canet Solà, Helena Hanna Juht, Maximilian Schich. 2023. A Framework for the Analysis of Historical Newsreels | [preprint](https://doi.org/10.31235/osf.io/a4xsp)
* Mehmet Burak Yilmaz, Elen Lotman, Andres Karjus, Pia Tikka. 2023. An embodiment of the cinematographer: emotional and perceptual responses to different camera movement techniques. Frontiers Neuroscience | [open access](https://doi.org/10.3389/fnins.2023.1160843) | [data](https://zenodo.org/record/7979713)
* Vejune Zemaityte, Andres Karjus, Ulrike Rohn, Maximilian Schich, Indrek Ibrus. 2023. Quantifying the global film festival circuit: Networks, diversity, and public value creation | [preprint](https://doi.org/10.31235/osf.io/g9w4b) | [code & data](https://andreskarjus.github.io/cinandofestivals)

2022

* Indrek Ibrus, Andres Karjus, Vejune Zemaityte, Ulrike Rohn, Maximilian Schich. 2022. Quantifying public value creation by public service media using big programming data | [preprint](https://osf.io/preprints/socarxiv/hysma/)
* Mark Mets, Andres Karjus, Indrek Ibrus, Maximilian Schich. 2023. Automated stance detection in complex topics and small languages: the challenging case of immigration in polarizing news media | [preprint](https://arxiv.org/abs/2305.13047)
* Andres Karjus, Christine Cuskley. 2022. Evolving linguistic divergence in socio-political polarities. Proceedings of the Joint Conference on Language Evolution (JCoLE). | [proceedings pdf](http://www.evolang.org/jcole2022_proceedings/jcole2022_proceedings.pdf)

2021

@@

2020

* Andres Karjus, 2020. Competition, selection and communicative need in language change. PhD thesis, University of Edinburgh | [pdf](https://andreskarjus.github.io/thesis/Andres_Karjus_phd_thesis.pdf) | [1-page non-technical summary](https://andreskarjus.github.io/thesis/english_summary.pdf) | [eestikeelne lühikokkuvõte](https://andreskarjus.github.io/thesis/doktoritoo_ylevaade_AK.pdf)
*  Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Communicative need modulates competition in language change | [preprint](https://arxiv.org/abs/2006.09277)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Challenges in detecting evolutionary forces in language change using diachronic corpora. Glossa: a journal of general linguistics, 5(1), p.45. | [open access](http://doi.org/10.5334/gjgl.909) | [code](https://github.com/andreskarjus/wfsim_fit)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2020. Quantifying the dynamics of topical fluctuations in language. Language Dynamics and Change 10(1), 86-125 | [open access](https://doi.org/10.1163/22105832-01001200) | [code](https://github.com/andreskarjus/topical_cultural_advection_model)

2018

* Andres Karjus, Martin Ehala, 2018. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change, 8, pp. 219-252 | [journal link](https://doi.org/10.1163/22105832-00802004) | [open postprint](Karjus_Ehala_2018_Testing_an_agent-based_model_of_language_choice_LDCpostprint.pdf) | <a href="karjus_testing_2018.bib" target="_blank">bib</a> | sociolinguistic [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): 1000 respondents, 200 questions
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract, full paper above] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>

2017

@@

- Talk at the Tallinn University School of Humanities seminar on AI & humanities research
- Seminar talk at the Change is Key project at the University of Gothenburg.
- Talk on LLMs for studying texts and change at "Assessing and measuring systems change" workshop | [slides](https://andreskarjus.github.io/talks/Tartu_LLMtalk.pdf)
- Poster on film festival research at Netsci 2023, Vienna
- Presentation at the Bibliotheca Herziana workshop on computational approaches to art
- Talk "Programming, data visualization & AI  for academic audiences across institutions and disciplines: lessons learned", at the Cross-university collaboration in Digital Humanities & Social Science (DHSS) and Digital Humanities & Cultural Heritage (DHCH) Education workshop of the DHNB2023 conference | [slides](https://datafigure.eu/reports/AndresKarjusDatafigure_CrossUniversityDHworkshop.pdf)
- Talk "Exploring Estonian Public Television Production 2004-2020 Using Big Programming Data" at the 8th Estonian Digital Humanities [Conference](https://dh.org.ee/events/dhe2022) (05.10.2022)
- Poster "Evolving Linguistic Divergence in Socio-Political Polarities", at the [JCoLe](https://sites.google.com/view/joint-conf-language-evolution) Joint Conference on Language Evolution in Kanazawa, Japan (August 2022)
- Poster "Linguistic divergence in American English along socio-political polarities", the [IC2S2](https://www.ic2s2.org/) Computational Social Science Conference (20.07.2022) | [pdf](https://andreskarjus.github.io/talks/ic2s2/IC2S2_poster_Linguistic_divergence_in_American_English_along_socio-political_polarities.pdf)
- Seminar talk at the Poncelet laboratory in Moscow (November 2021)
- [Conference on Complex Systems 2021](https://ccs2021.univ-lyon1.fr/) (October 2021) | [slides](https://andreskarjus.github.io/talks/ccs2021/Karjus_CCS2021_Capturing_aesthetic_complexity.pdf)
- [Protolang 7](https://blogs.phil.hhu.de/protolang7/programme/) (September 2021) | [slides](https://andreskarjus.github.io/talks/andres_karjus_protolang7.pdf)
- [Culture Conference 2021](https://culture-conference.com/) | [Poster on aesthetic complexity](https://andreskarjus.github.io/talks/cultureconference2021/)
- [TÜling](https://www.keel.ut.ee/et/tuling) (April 2021)
- [Colloquium for Computational Linguistics and Linguistics in Stuttgart](https://www.ims.uni-stuttgart.de/en/research/colloquium/) | [Slides](https://andreskarjus.github.io/talks/stuttgart2021/andres_karjus_lexical_dynamics_slides_stuttgart2021.pdf) | [Recording](https://www.ims.uni-stuttgart.de/documents/aktuelles/veranstaltungen/210111-Karjus-Talk-Colloq-Recording.mp4)
- [RUSE 2019](http://rusesymposium.org.uk/). Slides [here](https://andreskarjus.github.io/ruse2019).
- [CL2019](http://www.cl2019.org/). Slides [here](https://andreskarjus.github.io/cl2019talk.html).
- [Culture Conference 2019](https://culture-conference.com/programme-2019/), Poster [here](https://andreskarjus.github.io/lexcom_poster).
- [Inaugural ISLE workshop](https://www.comparativelinguistics.uzh.ch/en/events/ISLE-inaugural-workshop.html) | _Modelling lexical interactions in diachronic corpora_ |  [poster](https://andreskarjus.github.io/isle_poster)
- University of Edinburgh [Centre for Language Evolution](http://www.lel.ed.ac.uk/cle/) seminar series | _Challenges in detecting evolutionary forces in language change using diachronic corpora_ | [slides](https://andreskarjus.github.io/evoforces_cletalk/slides.html) | [code](https://github.com/andreskarjus/wfsim_fit)
- Corpus Linguistics in Scotland Network Meeting, _Topical Fluctuations and Lexical Interactions in Diachronic Corpora_

@@

I also teach workshops as an instructor in the private sector. 
Some of these have been standalone events, some have been part of conferences, summer schools or academic retreats. 
Feel free to get in touch if you are interested in talking about organizing a workshop on anything related to data science and statistics, artificial intelligence, data visualization, R, corpus linguistics, digital humanities, etc. <br>
For more details and contact, head over to [datafigure.eu](https://datafigure.eu)

@@

Academic email: andres.karjus --at-- tlu.ee <br>
Business and workshop inquiries: kindly contact via the <a href=https://datafigure.eu/contact.html</a> email  
<br>
<br>
Twitter / X: <a href=https://twitter.com/AndresKarjus>twitter.com/AndresKarjus</a> <br>
Bluesky: <a href=https://bsky.app/profile/andreskarjus.bsky.social>bsky.app/profile/andreskarjus.bsky.social</a> <br>
Mastodon: <a href=https://mastodon.social/@andreskarjus>mastodon.social/@AndresKarjus</a>
<br>
LinkedIn: <a href="https://www.linkedin.com/in/andreskarjus/">www.linkedin.com/in/andreskarjus/</a> <br> 
</div>

### 2023-11-12 news.Rmd
news

>>>>

- Now out in Corpus Linguistics and Linguistic Theory: "Reliable identification of selection mechanisms in language change" | [Open access](https://doi.org/10.1515/cllt-2023-0064)

- Mentored at the EdTech [Hackathon](https://eventornado.com/event/edtech-hack-2023?s=1#home) 2023 in Tallinn

- New preprint [Machine-assisted mixed methods: augmenting humanities and social sciences with artificial intelligence](https://arxiv.org/abs/2309.14379) (24.09.2023)
